Attrition in Massive Open Online Courses: Inequalities, Dropout Reasons, and Psychological Factors

### 2014-11-01 paper_v2.Rnw
small change to submitted draft. added new draft with updated fig

A number of significant and substantial individual differences emerged, some indicative of potential inequalities: Women were 10 to 15\% points less likely than men to persist with lectures and assessments. Women are also less likely to score a grade above the 60$^\text{th}$ percentile than men. Learners in Africa, Asia, and South America were 20 to 60\% points less likely to persist in video watching and achieved lower grades than learners in North America. Learners in Europe exhibited 10 to 30\% points higher persistence with lectures and assessemnts than those in North America. While learners in Asia showed lower persistence with videos, they were more likely to attempt more assessments than learners from North America.

@@

<<s1coefs, echo=FALSE, fig.env='figure*', fig.pos='ht', fig.height=6, fig.width=10, fig.cap='Coefficients of predictors (binary, except for normalized age) from logistic mixed-effects model predicting persistence and performance milestones (prop. videos watched, prop. assessments completed, and grade percentile). The baseline group were male learners in North America of average age who had not earned a bachelor\'s degree, without prior experience in the topic or other MOOCs, not intending to complete the course, and prepared to spend under two hours per week on it. An estimate of -20 indicates a 20\\% points lower probability of achieving a milestone.'>>=
library(ggplot2)
res = read.csv("pred_coefs.csv")
res$y = factor(res$y, levels=c(">30% videos", ">50% videos", ">80% videos", ">60% pctile grade", ">30% assessments", ">50% assessments", ">80% assessments", ">80% pctile grade"))
res$y = factor(res$y, labels=c("> 30% videos", "> 50% videos", "> 80% videos", "> 60th percentile grade", "> 30% assessments", "> 50% assessments", "> 80% assessments", "> 80th percentile grade"))
res$coef2 = factor(res$coef2, levels=rev(c("Intercept", "Gender:F", "Age (norm.)", "Educ.:Bachelor+", "Prior Exp.", "Intent:All", "Hours:2-4", "Hours:4+ ", "Prior MOOCs:1-3", "Prior MOOCs:4+ ", "Geo:Europe", "Geo:Oceania", "Geo:Africa", "Geo:Asia", "Geo:S.America")))

# This is what we submitted
ggplot(subset(res, coef2!="Intercept"), 
       aes(coef2, 100*est, ymin=100*(est-1.96*se), ymax=100*(est+1.96*se), color=factor(sign))) + 
  geom_hline(xintercept=0) + geom_pointrange() + facet_wrap(~y, ncol=4) + coord_flip() + 
  theme_minimal() + theme(legend.position="none",axis.title=element_text(vjust=0)) + 
  scale_color_manual(values=c('red','grey30','green3')) +
  labs(y="Predicted change in probability of achieving each `milestone` (95% CIs)", x="")

# This is what it should be
# ggplot(subset(res, coef2!="Intercept"), 
#   aes(coef2, 100*(exp(est)-1), ymin=100*(exp(est-1.96*se)-1), ymax=100*(exp(est+1.96*se)-1), color=factor(sign))) + 
#   geom_hline(yintercept=0) + geom_pointrange() + facet_wrap(~y, ncol=4) + coord_flip() + 
#   theme_minimal() + theme(legend.position="none",axis.title=element_text(vjust=0)) + 
#   scale_color_manual(values=c('red','grey30','green3')) +
#   labs(y="Predicted change in probability of achieving each `milestone` (95% CIs)", x="")

@

>>>>

A number of significant and substantial individual differences emerged, some indicative of potential inequalities: Women were 10 to 15\% less likely than men to persist with lectures and assessments. Women are also less likely to score a grade above the 60$^\text{th}$ percentile than men. Learners in Africa, Asia, and South America were 20 to 50\% less likely to persist in video watching and tended to achieve lower grades than learners in North America. In contrast, learners in Europe exhibited 20 to 50\% higher persistence with lectures and assessemnts than learners in North America. While learners in Asia exhibited lower persistence with videos, they were more likely to attempt more assessments than learners from North America.

@@

<<s1coefs, echo=FALSE, fig.env='figure*', fig.pos='ht', fig.height=6, fig.width=10, fig.cap='Coefficients of predictors (binary, except for normalized age) from logistic mixed-effects model predicting persistence and performance milestones (prop. videos watched, prop. assessments completed, and grade percentile). The baseline group were male learners in North America of average age who had not earned a bachelor\'s degree, without prior experience in the topic or other MOOCs, not intending to complete the course, and prepared to spend under two hours per week on it. An estimate of -20 indicates a 20\\% lower probability of achieving a milestone.'>>=
library(ggplot2)
res = read.csv("pred_coefs.csv")
res$y = factor(res$y, levels=c(">30% videos", ">50% videos", ">80% videos", ">60% pctile grade", ">30% assessments", ">50% assessments", ">80% assessments", ">80% pctile grade"))
res$y = factor(res$y, labels=c("> 30% videos", "> 50% videos", "> 80% videos", "> 60th percentile grade", "> 30% assessments", "> 50% assessments", "> 80% assessments", "> 80th percentile grade"))
res$coef2 = factor(res$coef2, levels=rev(c("Intercept", "Gender:F", "Age (norm.)", "Educ.:Bachelor+", "Prior Exp.", "Intent:All", "Hours:2-4", "Hours:4+ ", "Prior MOOCs:1-3", "Prior MOOCs:4+ ", "Geo:Europe", "Geo:Oceania", "Geo:Africa", "Geo:Asia", "Geo:S.America")))

# This is what we submitted
# ggplot(subset(res, coef2!="Intercept"), 
#        aes(coef2, 100*est, ymin=100*(est-1.96*se), ymax=100*(est+1.96*se), color=factor(sign))) + 
#   geom_hline(xintercept=0) + geom_pointrange() + facet_wrap(~y, ncol=4) + coord_flip() + 
#   theme_minimal() + theme(legend.position="none",axis.title=element_text(vjust=0)) + 
#   scale_color_manual(values=c('red','grey30','green3')) +
#   labs(y="Predicted change in probability of achieving each `milestone` (95% CIs)", x="")

# This is what it should be
ggplot(subset(res, coef2!="Intercept"), 
  aes(coef2, 100*(exp(est)-1), ymin=100*(exp(est-1.96*se)-1), ymax=100*(exp(est+1.96*se)-1), color=factor(sign))) + 
  geom_hline(yintercept=0) + geom_pointrange() + facet_wrap(~y, ncol=4) + coord_flip() + 
  theme_minimal() + theme(legend.position="none",axis.title=element_text(vjust=0)) + 
  scale_color_manual(values=c('red','grey30','green3')) +
  scale_y_continuous(breaks=c(-60,-40,-20,0,20,40,60)) +
  labs(y="Predicted change in probability of achieving each `milestone` (95% CIs)", x="")
coefs = subset(res, coef2=="Intercept", select=c("est","y"))  # should report these in the text of reference
coefs$exp_est = exp(coefs$est) # estimates seem high
@
